---
title: "Assignment 2 - Functions, iteration and debugging"
author: "Daniyar Imanaliev"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<!-- Do not forget to input your Github username in the YAML configuration up there --> 

Daniyar2822

```{r, include = T}
library(tibble)
library(tidyverse)
library(dplyr)
library(eeptools)
library(unvotes)
library(countrycode)
```

<br>

### Task 1 - Fun with functions

a) Program a function `ultimate_answer()` that always returns the number 42 regardless of which input is provided, and show that it works providing three function calls that test different input types!

```{r}
ultimate_answer <- function(input) {
  return(42)
}
ultimate_answer(34)
ultimate_answer('hello')
ultimate_answer(TRUE)
```

b) Write an R function called `color_guessing_game()` that allows the player to guess one of four colors (red, yellow, green, blue). The game will randomly select one of the colors and prompt the player to guess it. If the player's guess matches the selected color, they win, otherwise, they lose. *Hint:* You will need the function `readline()` to enable interactive communication between the game and the player.

```{r}
game <- function() {
  #creating sample of colors using sample function
  colors <- sample(c("red", "yellow", "green", "blue"), size = 1)
  #creating interactive part of function for receiving input from the user
  cat("Hello! Choose the color (red, yellow, green, or blue): ")
  player <- tolower(readline())
  if (player == colors) {
    cat("Congratulations! You won")
  } else {
    cat("Sorry, you lost") 
  }
}
  game()

```

<br>

***

### Task 2 - Functional programming in practice

The `unvotes` package (hosted both on [CRAN](https://cran.r-project.org/web/packages/unvotes/index.html) and [GitHub](https://github.com/dgrtwo/unvotes)) provides data on the voting history of countries in the [United Nations General Assembly](http://www.un.org/en/ga/), along with information such as date, description, and topics for each vote. The package contains three datasets: `un_votes`, providing the history of each country's votes, `un_roll_calls`, providing information on each roll call vote, and `un_roll_call_issues`, providing issue (topic) classifications of roll call votes. Check out the [database tutorial](https://github.com/dgrtwo/unvotes) if you want to make yourself familiar with the data structure. Then, work on the following tasks.

a) Calculate how often, on average, Germany agreed with the US on a vote in the UN general assembly since 1990. Votes where voting information for one of the countries is missing should be discarded.

```{r}
# combining 'un_roll_calls' dataset and 'un_votes' dataset
un <- full_join(un_roll_calls, un_votes, by = "rcid", relationship = "many-to-many")

# creating specific dataframe only for US and Germany data, filtering data from 1990 removing missed ones
de_us <- un %>%
  filter(date >= '1990-01-01', !is.na(vote) & !is.na(country)) %>%
  # it is important to note that before 03 October 1990 (date of unification of Germany), the name of Federal Republic of Germany was used, 
  # accordingly for purposes of this task, name of 'Federal Republic of Germany' shall be also used
  mutate(country_code = ifelse(country == "Federal Republic of Germany", "DE", country_code)) %>%
  filter(country %in% c("Germany", "Federal Republic of Germany", "United States")) %>%
  #removing redundant objects, remaining only rcid, country_code and vote
  select(rcid, country_code, vote) %>%
  #creating object for US and DE votes
  pivot_wider(names_from = country_code, values_from = vote) %>%
  #removing missed values
  drop_na()
#calculating average 
agr_rate <- mean(de_us$DE == de_us$US) #it is important to note that abstain votes also included to calculations, since abstain voting also shows political position of UN members
cat("On average, from 1990 Germany agreed with the US in ", round(agr_rate*100, 2), "%", sep = "")
```

<br>

b) Now, create a function, `votes_agreement_calculator()`, that takes country identifiers as well as a `year_min` argument as inputs and that returns the share of agreement in voting between any two specified countries as numeric value, for a time period specified with year >= `year_min`. The function should take the necessary data frames directly from the `unvotes` package. Then, test the function by computing the agreement rates for (a) the United States and Russia for votes cast in 2000 and later and (b) France and the UK for votes cast in 2000 and later!

```{r}
#creating function using our previous code
votes_agreement_calculator <- function(code = c('DE', 'US'), year_min = 1990) {
  un <- full_join(un_roll_calls, un_votes, by = "rcid", relationship = "many-to-many") %>%
    select(rcid, date, country_code, vote) %>%
    filter(!is.na(vote) & !is.na(country_code)) %>%
    filter(date >= paste(year_min, '-01-01', sep = "")) #using only year as required
  
  if (sum(un$country_code %>% unique() %in% code) <= 1) {
    return(0) #it is necessary to check if input contains correct code
  } else {
    un_members <- un %>% filter(country_code %in% code) %>%
      pivot_wider(names_from = country_code, values_from = vote) %>%
      drop_na()
    agr_rate <- mean(un_members[[code[1]]] == un_members[[code[2]]])
    return(round(agr_rate * 100, 2))
  }
}
```


```{r}
#checking our code. Please note that result will be different from previous task, since it was intentionally ignored special cases like unification of Germany

votes_agreement_calculator(code=c('US','DE'),year_min = 1990)
```


```{r}
#calculating (a) the United States and Russia for votes cast in 2000 and later

votes_agreement_calculator(code=c('RU','US'),year_min = 2000)

```
```{r}
#calculating (b) France and the UK for votes cast in 2000 and later

votes_agreement_calculator(code=c('FR','GB'),year_min = 2000)

```


<br>

c) Using `purrr` functionality, find out which three countries on average agreed with the US the most from the year 2000 on!

```{r, eval = TRUE}
#creating object which excludes US as well as missed value
not_us <- un$country_code %>% na.exclude() %>% unique() %>% .[-which(. =="US")]
#with usage map_dbl function calculating every UN members on average percent of mutual agreement during UN voting in ascending order
top_allies <- not_us %>% 
  map_dbl(~(votes_agreement_calculator(code = c(.x,"US"),year_min = 2000))) %>% 
  order(decreasing = T)
# finding top three UN members
top_three_allies <- top_allies[c(1:3)]
cat("There are the following countries agreed with US the most from the year 2000:", countrycode(not_us[top_three_allies], origin = 'iso2c', destination = 'country.name'))


```

<br>

***

### Task 3 - Debugging code

The following code snippet contains various bugs. Flag them and suggest a fix by adding a comment in the respective line. Example:

```{r, eval = FALSE}
library(Tidyverse) # BUG: typo in library(Tidyverse); should be library(tidyverse) instead
```

```{r, eval = FALSE}
# load packages
library(tidyverse)
library(countrycode)
library(Unvotes) # BUG: typo in library(Unvotes); should be library(unvotes)
# get continents data
continents <- countrycode:codelist %>% # BUG: should be double colons (::) instead of single one
  select(continent, iso2c) %>%
  rename(country_code == iso2c)
un_votes <- left_join(x = un_votes, y = continents, by = country_code)
# get data on European UN votes
eu_un_votes <- left_join(un_votes, un_roll_calls, by = "rcid") %>%
  left_join(., un_roll_call_issues, by = "rcid") %>% 
  filter(continent == "Europe", 
         date >= "1991-01-01" & date <= "2021-12-31") %>% 
  drop_na(short_name)
# encode Europe's voting behavior numerically
eu_un_votes <- eu_un_votes %>% 
  mutate(yes_vt = if_else(vote == "yes", 1, 0)), #BUG: should only one parenthesis instead of two
         no_vt = if_else(vote == "no", 1, 0),
         abstention = if_else(vote == "abstain", 1, 0))
# list top 10 most disagreed with UN resolutions
hated_un <- eu_un_votes %>%
  group_by(unres) #BUG : pipe operator '%>%' missed
  summarise(sh_disagree_with = sum(no_vt,na.rm = T) / sum(c(yes_vt,no_vt, abstention), na.rm = T)) %>% 
  arrange(desc(sh_disagree_with)) %>% 
  head(10)
hated_un
``` 


